Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-29 aistats.tex
Introduce gamma in appendix A

\usepackage{array}

@@

In this section, we prove the results mentioned in \S\ref{ssec:gaussian-variance},
and add some context and experimental observations.
Let us recall the form of the divergence with different parametrizations.
For a Gaussian with known mean $0$ and covariance $\sigma^2$, 
natural and mean parameters $\theta, \m$;
\begin{align}
	\KL( \sigma_*^2 ; \sigma_n^2 ) 
	&= \half \left ( \frac{\sigma_*^2}{ \sigma_n^2} - 1 - \log \frac{\sigma_*^2}{ \sigma_n^2} \right)
	= \phi\paren{\frac{\sigma_*^2}{ \sigma_n^2}},
	\\
	\bregmanconj(\m_*; \m_n) 
	&= \half \left ( \frac{\m_*}{ \m_n} - 1 - \log  \frac{\m_*}{ \m_n} \right) 
	= \phi\paren{\frac{\m_*}{ \m_n}},
	\\
	\bregman( \nat_n; \nat_* ) 
	&=  \half \left ( \frac{ \nat_n}{\nat_*} - 1 - \log  \frac{ \nat_n}{\nat_*} \right)
	= \phi\paren{\frac{ \nat_n}{\nat_*}},
\end{align}
where all these lines are equal, and $\phi$ measures the discrepancy between the ratio $\frac{ \nat_n}{\nat_*} =  \frac{\m_*}{ \m_n}  $ and $1$ via the function
\begin{align}
	\phi(z) := \half (z - 1 - \log(z)),
\end{align}
illustrated in orange in Figure~\ref{fig:phi}.
To upper bound for the MAP, due to the difficulty of finding a closed form for the expectation of the logarithm, we focus on the symmetrized Bregman instead
\alignn{
	\bregmanconj( \mu_*; \mu_n) 
	\leq \cS_{\conj} (\mu_*, \mu_n ) 
	:= \bregmanconj( \mu_*; \mu_n)  + \bregmanconj( \mu_n; \mu_*) 
	= \half \paren{\frac{ \mu_*}{\mu_n} -1 +\frac{ \mu_n}{\mu_*} - 1 },
	\label{eq:symmetrized_bregman}
}
which can also be expressed with $\phi$ as follows:
\aligns{
	\label{eq:log_bound}
	\phi(z) \leq \phi(z) + \phi\paren{\inv{z}}
	&=  \half \paren{z -1  + \inv{z} -1} 
	= \frac{(z-1)^2}{z} \; .
}
See the grey upper bound in Figure~\ref{fig:phi}.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.4\textwidth]{phi.pdf}
	\caption{%
	$\phi(z)$ is the Bregman divergence induced by $-\log(z)$. 
	It is a barrier near $0$, and admits an upper bound (in grey) despite being poorly approximated by quadratics.
	}
	\label{fig:phi}
\end{figure}

>>>>

\usepackage{array}
\usepackage{wrapfig}

@@

In this section, we prove the results mentioned in \S\ref{ssec:gaussian-variance},
and add some context and experimental observations.
As mentioned in the main text, the centered gaussian $\cN(0,\sigma^2)$ has sufficient statistic $x^2$ following a Gamma distribution $\Gamma\paren{\half, \inv{2 \sigma^2}}$. 
In this section we derive results for samples  from a general gamma distribution $X \sim\Gamma(\alpha,\beta)$ with known shape parameter $\alpha$, but unknown rate parameter $\beta$.
Results for the Gaussian follow by taking $\alpha=\half$.

The density of $\Gamma(\alpha,\beta)$ reads
\alignn{
	p(x) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x} \; .
}
When $\alpha$ is known, it can be cast as an exponential family~\eqref{eq:def_expfamily} with sufficient statistic $T(x)=x$, domain $\cX = \real_+$ and base measure $\nu(x)\propto x^{\alpha-1}$. Then the natural parameter is $\nat = -\beta < 0$ and the log-partition function is
\alignn{
	\logpart (\nat) = -\alpha \log(-\nat)  + \log\Gamma(\alpha) \; .
}
From there we find that the mean parameter is $\mu = \frac{\alpha}{-\nat} >0$ and the entropy has the same form as the log-partition $\conj(\mu) = - \alpha \log(\mu) + \cst$.
This means that the primal and dual Bregman divergences have the same form as well
\begin{align}
	\bregmanconj(\m_*; \m_n) 
	&= \alpha \left ( \frac{\m_*}{ \m_n} - 1 - \log  \frac{\m_*}{ \m_n} \right) 
	= \alpha \phi\paren{\frac{\m_*}{ \m_n}},
	\\
	\bregman( \nat_n; \nat_* ) 
	&=  \alpha \left ( \frac{ \nat_n}{\nat_*} - 1 - \log  \frac{ \nat_n}{\nat_*} \right)
	= \alpha \phi\paren{\frac{ \nat_n}{\nat_*}},
\end{align}
\begin{wrapfigure}[14]{r}{0.4\textwidth}
	\centering
	\includegraphics[width=.38\textwidth]{phi.pdf}
	\caption{%
	Illustration of $\phi(z)$ (orange) and its upper bound $\phi(z) + \phi\paren{z^{-1}}$ (grey).
	They both are barriers near $0$.
	}
	\label{fig:phi}
\end{wrapfigure}
where these 2 lines are equal, and $\phi$ measures the discrepancy between the ratio $\frac{ \nat_n}{\nat_*} =  \frac{\m_*}{ \m_n}  $ and $1$ via the function
\begin{align}
	\phi(z) := z - 1 - \log(z),
\end{align}
illustrated in orange in Figure~\ref{fig:phi}.
To upper bound for the MAP, due to the difficulty of finding a closed form for the expectation of the logarithm, we focus on the symmetrized Bregman instead
\alignn{
	\bregmanconj( \mu_*; \mu_n) 
	\leq \cS_{\conj} (\mu_*, \mu_n )
	&:= \bregmanconj( \mu_*; \mu_n)  + \bregmanconj( \mu_n; \mu_*) \nonumber\\
	&=  \alpha \paren{\frac{ \mu_*}{\mu_n} -1 +\frac{ \mu_n}{\mu_*} - 1 },
	\label{eq:symmetrized_bregman}
}
which can also be expressed with $\phi$ as follows:
\aligns{
	\label{eq:log_bound}
	\phi(z) \leq \phi(z) + \phi\paren{z^{-1}}
	&=  z -1  + \inv{z} -1
	= \frac{(z-1)^2}{z} \; ,
}
and as illustrated by the grey upper bound in Figure~\ref{fig:phi}.
